---
title: "Deseq2_Oli_analysis"
author: "Гринберг Михаил"
date: "2024-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tximport)
library(DESeq2)
library(ggplot2)
library(ggalt)
library(dplyr)
library(tibble)
library(topGO)
library(rrvgo)
library(viridisLite)
library(pheatmap)

library(ggpubr)
library(ggVennDiagram)
library(UpSetR)
library(tibble)
library(stringr)

source("/Volumes/oli/RESULTS/scripts/07_GODeseq/deseq2_utils.R")
```

## Imports + sample names


```{r imports, message=FALSE, warning=FALSE}
annotation <- read.csv("/Volumes/oli/RESULTS/data/04_annotation/eggnog/oli_annot.csv", sep = '\t')

oli_geneID2GO_file <- "/Volumes/oli/RESULTS/data/04_annotation/eggnog/oli_gene2GO.csv"
oli_geneID2GO <- readMappings(file = oli_geneID2GO_file, sep = '\t')
gene_names <- names(oli_geneID2GO)

gene2trans <- read.table(file = "/Volumes/oli/RESULTS/data/01_assembly/gene2trans.txt")
gene2trans <- gene2trans[,c(2,2)]

samples_names <- c("OL1.out", "OL2.out", "OL3.out","OL5.out",
                     "OL6.out","OL7.out","OL9.out","OL10.out",
                     "OL11.out","OL14.out", "OL15.out", "OL16.out", 
                     "OL18.out","OL19.out","OL20.out")

files <- file.path("/Volumes/oli/RESULTS/data/02_salmon/salmon", samples_names, "filtered_quant.sf")
names(files) <- c("egg_1", "egg_2", "egg_3", 
                                     "blastula_1", "blastula_2", "blastula_3", 
                                     "gastrula_1", "gastrula_2", "gastrula_3",  
                                     "trochophore_1", "trochophore_2", "trochophore_3",
                                     "adult_1", "adult_2", "adult_3")

txi<- tximport(files, type = "salmon", tx2gene = gene2trans, countsFromAbundance="lengthScaledTPM")
```

## Condition and sampleTable creation

```{r cond, message=FALSE, warning=FALSE}
cond <- data.frame(read.csv("/Volumes/oli/RESULTS/scripts/07_GODeseq/deseq2_cond.txt", 
                            header=T, sep = '\t', row.names = 1))
cond$Condition <- factor(cond$Condition, levels = unique(cond$Condition))
sampleTable <- data.frame(condition = cond$Condition)
rownames(sampleTable) <- colnames(txi$counts)
```

## Deseq2 object creation

```{r deseq2, message=FALSE, warning=FALSE}
dds <- DESeqDataSetFromTximport(txi, sampleTable, ~condition)
```

## Creating table with normalised counts 

```{r norm counts}
dds <- estimateSizeFactors(dds)
normalized_counts <- counts(dds, normalized=TRUE)
normalized_counts <- as.data.frame(normalized_counts)
normalized_counts <- rownames_to_column(normalized_counts, var = "gene")
write.table(normalized_counts, file="normalized_counts.txt", sep="\t", quote=F, col.names=NA)
```

## Stages clustering

```{r pca, message=FALSE, warning=FALSE}
vst_dds <- vst(dds)
counts.norm <- assay(vst_dds)

plotPCA(vst_dds,intgroup=c("condition")) + 
  geom_encircle(aes(x = PC1, y = PC2, color = condition)) 
```


```{r sample clustering, message=FALSE, warning=FALSE}
sampleDists <- dist(t(assay(vst_dds)))
sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- paste(vst_dds$condition, vst_dds$type)

pheatmap(sampleDistMatrix,
         annotation_col = sampleTable,
         clustering_distance_rows = sampleDists,
         clustering_distance_cols = sampleDists, 
         color = viridis(100),
         show_colnames = FALSE)
```

## All vs egg

Comparasion of all stages to a starting point - an egg:

### DESeq
```{r dds_vs_egg, message=FALSE, warning=FALSE}
dds$condition = relevel(dds$condition, "egg")
dds_vs_egg <- DESeq(dds)
```

### Plots

```{r deseq2_results_vs_egg}
contrast1 = c("condition", "blastula","egg")
contrast2 = c("condition", "gastrula","egg")
contrast3 = c("condition", "trochophore","egg")
contrast4 = c("condition", "adult","egg")

blast_vs_egg <- results(dds_vs_egg, contrast = contrast1)
gastr_vs_egg <- results(dds_vs_egg, contrast = contrast2)
troch_vs_egg <- results(dds_vs_egg, contrast = contrast3)
adult_vs_egg <- results(dds_vs_egg, contrast = contrast4)

blast_vs_egg <- classify_DEGs(blast_vs_egg)
gastr_vs_egg <- classify_DEGs(gastr_vs_egg)
troch_vs_egg <- classify_DEGs(troch_vs_egg)
adult_vs_egg <- classify_DEGs(adult_vs_egg)
```

```{r volcano_all_vs_egg, message=FALSE, warning=FALSE}
volcano_plot(blast_vs_egg$all, contrast1)
volcano_plot(gastr_vs_egg$all, contrast2)
volcano_plot(troch_vs_egg$all, contrast3)
volcano_plot(adult_vs_egg$all, contrast4)
```
Venn diagramm for all DEGs:

```{r venn_all_vs_egg}
blast_vs_egg_sig <- blast_vs_egg$all[blast_vs_egg$all$DEGs != "Non-DEG", ]
gastr_vs_egg_sig <- gastr_vs_egg$all[gastr_vs_egg$all$DEGs != "Non-DEG", ]
troch_vs_egg_sig <- troch_vs_egg$all[troch_vs_egg$all$DEGs != "Non-DEG", ]
adult_vs_egg_sig <- adult_vs_egg$all[adult_vs_egg$all$DEGs != "Non-DEG", ]

blast_vs_egg_sig_names <- blast_vs_egg_sig$gene
gastr_vs_egg_sig_names <- gastr_vs_egg_sig$gene
troch_vs_egg_sig_names <- troch_vs_egg_sig$gene
adult_vs_egg_sig_names <- adult_vs_egg_sig$gene

ggVennDiagram(list(blast_vs_egg_sig_names, gastr_vs_egg_sig_names, troch_vs_egg_sig_names, adult_vs_egg_sig_names), 
              category.names = c("blastula" , "gastrula", "trochophore", "adult"),
              label_alpha=0.8, cat.cex = 1,
              cat.fontface = "bold",
              cat.default.pos = "outer",
              cat.dist = c(0.055, 0.055, 0.1, 0.1)) + 
  
  scale_fill_gradient(low="blue",high = "red")
```

UpsetR plots for upregulated and downregulated DEGs:

```{r upsetr_all_vs_egg, message=FALSE, warning=FALSE}
upregulated <- list(blastula = blast_vs_egg$up$gene, 
                  gastrula = gastr_vs_egg$up$gene, 
                  trochophore = troch_vs_egg$up$gene,
                  adult = adult_vs_egg$up$gene)

downregulated <- list(blastula = blast_vs_egg$down$gene, 
                  gastrula = gastr_vs_egg$down$gene, 
                  trochophore = troch_vs_egg$down$gene,
                  adult = adult_vs_egg$down$gene)

upset(fromList(upregulated), order.by = "freq", sets = names(upregulated), 
      mainbar.y.label = "Upregulated genes intersections")

upset(fromList(downregulated), order.by = "freq", sets = names(downregulated), 
      mainbar.y.label = "Downregulated genes intersections")
```

### Blastula analysis

```{r blast_vs_egg_up_GO, message=FALSE, warning=FALSE}
blast_vs_egg_up_GO <- run_GO_analysis(blast_vs_egg$up, oli_geneID2GO)
blast_vs_egg_up_rTerms_uniq <- blast_vs_egg_up_GO$reducedTerms_uniq
blast_vs_egg_up_rTerms <- blast_vs_egg_up_GO$reducedTerms
blast_vs_egg_up_allTerms <- blast_vs_egg_up_GO$GO_subset
```

```{r blast_vs_egg_up_plots}
treemapPlot(blast_vs_egg_up_rTerms, size = "score")
plot_GO_hist(blast_vs_egg_up_rTerms_uniq, "blastula vs egg: Upregulated")
```

```{r GO:0048852_heatmap}
dienm_genes <- get_term_genes(blast_vs_egg$up, "GO:0048852", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(dienm_genes, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "diencephalon morphogenesis (GO:0048852)")
```

```{r GO:0021532_heatmap}
neut <- get_term_genes(blast_vs_egg$up, "GO:0021532", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(neut, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Neural tube formation (GO:0021532)")
```

```{r GO:0030326+GO:0035113 heatmap}
animal_organ_morph <- rbind(
                            get_term_genes(blast_vs_egg$up, "GO:0035113", oli_geneID2GO, annotation), 
#embryonic appendage morphogenesis
                            get_term_genes(blast_vs_egg$up, "GO:0030326", oli_geneID2GO, annotation)) 
#embryonic limb morphogenesis

#get_term_genes(blast_vs_egg$up, "GO:0060411"),
#get_term_genes(blast_vs_egg$up, "GO:0003151"),
#get_term_genes(blast_vs_egg$up, "GO:0060412"),
#get_term_genes(blast_vs_egg$up, "GO:0031290"),
#get_term_genes(blast_vs_egg$up, "GO:0060349"))

animal_organ_morph <- unique(animal_organ_morph)

zscore <- make_zscore_matrix(animal_organ_morph, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Animal organ morphogenesis (GO:0035113, GO:0030326)")

#CTNNB1 - beta-catenin
```

```{r blast_vs_egg_down_GO, message=FALSE, warning=FALSE}
blast_vs_egg_down_GO <- run_GO_analysis(blast_vs_egg$down, oli_geneID2GO)
blast_vs_egg_down_rTerms_uniq <- blast_vs_egg_down_GO$reducedTerms_uniq
blast_vs_egg_down_rTerms <- blast_vs_egg_down_GO$reducedTerms
blast_vs_egg_down_allTerms <- blast_vs_egg_down_GO$GO_subset
```

```{r blast_vs_egg_down_plots}
treemapPlot(blast_vs_egg_down_rTerms, size = "score")
plot_GO_hist(blast_vs_egg_down_rTerms_uniq, "blastula vs egg: Downregulated")
```
### Gastrula analysis

```{r gastr_vs_egg_up_GO, message=FALSE, warning=FALSE}
gastr_vs_egg_up_GO <- run_GO_analysis(gastr_vs_egg$up, oli_geneID2GO)
gastr_vs_egg_up_rTerms_uniq <- gastr_vs_egg_up_GO$reducedTerms_uniq
gastr_vs_egg_up_rTerms <- gastr_vs_egg_up_GO$reducedTerms
gastr_vs_egg_up_allTerms <- gastr_vs_egg_up_GO$GO_subset
```

```{r gastr_vs_egg_up_plots}
treemapPlot(gastr_vs_egg_up_rTerms, size = "score")
plot_GO_hist(gastr_vs_egg_up_rTerms_uniq, "gastrula vs egg: Upregulated")
```

```{r GO:0023019_heatmap}
signal_transduction <- get_term_genes(gastr_vs_egg$up, "GO:0023019", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(signal_transduction, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Signal transduction (GO:0023019)")
```

```{r GO:0031128_heatmap}
dev_ind <- get_term_genes(gastr_vs_egg$up, "GO:0031128", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(dev_ind, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Developmental induction (GO:0031128)")
```

```{r GO:0071697_heatmap}
ectodermal_placode <- get_term_genes(gastr_vs_egg$up, "GO:0071697", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(ectodermal_placode, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Ectodermal placode morphogenesis (GO:0071697)")
```

```{r dds_vs_blast, message=FALSE, warning=FALSE}
dds$condition = relevel(dds$condition, "blastula")
dds_vs_blast <- DESeq(dds)
```

```{r deseq_results_gastr_vs_blast, message=FALSE, warning=FALSE}
contrast = c("condition", "gastrula","blastula")
gastr_vs_blast <- results(dds_vs_blast, contrast = contrast)
gastr_vs_blast <- classify_DEGs(gastr_vs_blast)
volcano_plot(gastr_vs_blast$all, contrast)
```


```{r gastr_vs_blast_up_GO, message=FALSE, warning=FALSE}
gastr_vs_blast_up_GO<- run_GO_analysis(gastr_vs_blast$up, oli_geneID2GO)
gastr_vs_blast_up_rTerms_uniq <- gastr_vs_blast_up_GO$reducedTerms_uniq
gastr_vs_blast_up_rTerms <- gastr_vs_blast_up_GO$reducedTerms
gastr_vs_blast_up_allTerms <- gastr_vs_blast_up_GO$GO_subset
```

```{r gastr_vs_blast_up_plot}
treemapPlot(gastr_vs_blast_up_rTerms, size = "score")
plot_GO_hist(gastr_vs_blast_up_rTerms_uniq, "gastrula vs blastula: Upregulated")
```

```{r GO:0048505_heatmap}
cell_diff <- get_term_genes(gastr_vs_blast$up, "GO:0048505", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(cell_diff, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Regulation of timing of cell differantiation (GO:0023019)")
```

```{r GO:0003144_heatmap}
embryonic_heart <- get_term_genes(gastr_vs_blast$up, "GO:0003144", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(embryonic_heart, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Embryonic heart tube formation (GO:0003144)")
```

```{r GO:0048596_heatmap}
embr_eye <- get_term_genes(gastr_vs_blast$up, "GO:0048596", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(embr_eye, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Eye formation (GO:0048596)")
```

```{r GO:0048703_heatmap}
viscerocranium <- get_term_genes(gastr_vs_blast$up, "GO:0048703", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(viscerocranium, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Viscerocranium morphogenesis (GO:0048703)")
```

```{r GO:0001649_heatmap}
osteoblast <- get_term_genes(gastr_vs_blast$up, "GO:0001649", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(osteoblast, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Osteoblast differentiation (GO:0001649)")
```

```{r gastr_vs_blast_down_GO, message=FALSE, warning=FALSE}
gastr_vs_blast_down_GO<- run_GO_analysis(gastr_vs_blast$down, oli_geneID2GO)
gastr_vs_blast_down_rTerms_uniq <- gastr_vs_blast_up_GO$reducedTerms_uniq
gastr_vs_blast_down_rTerms <- gastr_vs_blast_up_GO$reducedTerms
gastr_vs_blast_down_allTerms <- gastr_vs_blast_up_GO$GO_subset
```

```{r gastr_vs_blast_down_plots}
treemapPlot(gastr_vs_blast_down_rTerms, size = "score")
plot_GO_hist(gastr_vs_blast_down_rTerms_uniq, "gastrula vs blastula: Downgulated")
```


```{r GO:0048596_heatmap2}
embr_eye <- get_term_genes(gastr_vs_blast$down, "GO:0048596", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(embr_eye, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Eye formation (GO:0048596)")
```

```{r tmp_heatmap}
#signal_transduction <- get_term_genes(gastr_vs_blast$down, "GO:0023019", oli_geneID2GO, annotation)
#embryonic_heart <- get_term_genes(gastr_vs_blast$down, "GO:0003144", oli_geneID2GO, annotation)
#cell_diff <- get_term_genes(gastr_vs_blast$down, "GO:0048505", oli_geneID2GO, annotation)

#each - 1 gene
```

### Trochophore analysis

```{r troch_vs_egg_up_GO, message=FALSE, warning=FALSE}
troch_vs_egg_up_GO <- run_GO_analysis(troch_vs_egg$up, oli_geneID2GO)
troch_vs_egg_up_rTerms_uniq <- troch_vs_egg_up_GO$reducedTerms_uniq
troch_vs_egg_up_rTerms <- troch_vs_egg_up_GO$reducedTerms
troch_vs_egg_up_allTerms <- troch_vs_egg_up_GO$GO_subset
```

```{r troch_vs_egg_up_plots}
treemapPlot(troch_vs_egg_up_rTerms, size = "score")
plot_GO_hist(troch_vs_egg_up_rTerms_uniq, "troch vs egg: Upregulated")
```

```{r GO:0061074_heatmap}
neural_retina <- get_term_genes(troch_vs_egg$up, "GO:0061074", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(neural_retina, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Neural retina formation (GO:0061074)")
```

```{r GO:0021871_heatmap}
forebrain <- get_term_genes(troch_vs_egg$up, "GO:0021871", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(forebrain, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Forebrain regionalisation (GO:0021871)")
```

```{r dds_vs_gastr, message=FALSE, warning=FALSE}
dds$condition = relevel(dds$condition, "gastrula")
dds_vs_gastr <- DESeq(dds)

contrast = c("condition", "trochophore","gastrula")
troch_vs_gastr <- results(dds_vs_gastr, contrast = contrast)
troch_vs_gastr <- classify_DEGs(troch_vs_gastr)

volcano_plot(troch_vs_gastr$all, contrast)
```

```{r troch_vs_gastr_up_GO, message=FALSE, warning=FALSE}
troch_vs_gastr_up_GO <- run_GO_analysis(troch_vs_gastr$up, oli_geneID2GO)
troch_vs_gastr_up_rTerms_uniq <- troch_vs_gastr_up_GO$reducedTerms_uniq
troch_vs_gastr_up_rTerms <- troch_vs_gastr_up_GO$reducedTerms
troch_vs_gastr_up_allTerms <- troch_vs_gastr_up_GO$GO_subset
```

```{r troch_vs_gastr_up_plots}
treemapPlot(troch_vs_gastr_up_rTerms, size = "score")
plot_GO_hist(troch_vs_gastr_up_rTerms_uniq, "troch vs gastrula: Upregulated")
```

```{r troch_vs_gastr_down_GO, message=FALSE, warning=FALSE}
troch_vs_gastr_down_GO <- run_GO_analysis(troch_vs_gastr$down, oli_geneID2GO)
troch_vs_gastr_down_rTerms_uniq <- troch_vs_gastr_down_GO$reducedTerms_uniq
troch_vs_gastr_down_rTerms <- troch_vs_gastr_down_GO$reducedTerms
troch_vs_gastr_down_allTerms <- troch_vs_gastr_down_GO$GO_subset
```

```{r troch_vs_gastr_down_plots}
treemapPlot(troch_vs_gastr_down_rTerms, size = "score")
plot_GO_hist(troch_vs_gastr_down_rTerms_uniq, "troch vs gastrula: Downregulated")
```

```{r GO:0007379_heatmap}
segment <- get_term_genes(troch_vs_gastr$down, "GO:0007379", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(segment, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Segment specification (GO:0007379)")
```

```{r GO:0045747_heatmap}
notch <- get_term_genes(troch_vs_gastr$down, "GO:0045747", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(notch, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Positive Notch regulation (GO:0045747)")
```

```{r GO:0001709_heatmap}
fate <- get_term_genes(troch_vs_gastr$down, "GO:0001709", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(fate, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Cell fate determination (GO:0001709)")
```

```{r GO:0048863_heatmap}
stemc <- get_term_genes(troch_vs_gastr$down, "GO:0048863", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(stemc, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Stem cell differentiation (GO:0048863)")
```

### Adult analysis

```{r adult_vs_egg_up_GO, message=FALSE, warning=FALSE}
adult_vs_egg_up_GO <- run_GO_analysis(adult_vs_egg$up, oli_geneID2GO)
adult_vs_egg_up_rTerms_uniq <- adult_vs_egg_up_GO$reducedTerms_uniq
adult_vs_egg_up_rTerms <- adult_vs_egg_up_GO$reducedTerms
adult_vs_egg_up_allTerms <- adult_vs_egg_up_GO$GO_subset
```

```{r adult_vs_egg_up_plots}
treemapPlot(adult_vs_egg_up_rTerms, size = "score")
plot_GO_hist(adult_vs_egg_up_allTerms, "adult vs egg: Upregulated")
```

```{r GO:0023019_heatmap2}
signal_transduction <- get_term_genes(adult_vs_egg$up, "GO:0023019", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(signal_transduction, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Signal transduction (GO:0023019)")
```

```{r GO:0021516_heatmap}
dors <- get_term_genes(adult_vs_egg$up, "GO:0021516", oli_geneID2GO, annotation)
zscore <- make_zscore_matrix(dors, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Dorsal spinal cord (GO:0021516)")
```

```{r dds_vs_troch, message=FALSE, warning=FALSE}
dds$condition = relevel(dds$condition, "trochophore")
dds_vs_troch <- DESeq(dds)

contrast = c("condition","adult", "trochophore")
adult_vs_troch <- results(dds_vs_troch, contrast = contrast)
adult_vs_troch <- classify_DEGs(adult_vs_troch)

volcano_plot(adult_vs_troch$all, contrast)
```

```{r adult_vs_troch_up_GO, message=FALSE, warning=FALSE}
adult_vs_troch_up_GO <- run_GO_analysis(adult_vs_troch$up, oli_geneID2GO)
adult_vs_troch_up_rTerms_uniq <- adult_vs_troch_up_GO$reducedTerms_uniq
adult_vs_troch_up_rTerms <- adult_vs_troch_up_GO$reducedTerms
adult_vs_troch_up_allTerms <- adult_vs_troch_up_GO$GO_subset
```

```{r adult_vs_troch_up_plot}
treemapPlot(adult_vs_troch_up_rTerms, size = "score")
plot_GO_hist(adult_vs_troch_up_allTerms, "adult vs troch: Upregulated")
```

```{r adult_vs_troch_down_GO, message=FALSE, warning=FALSE}
adult_vs_troch_down_GO <- run_GO_analysis(adult_vs_troch$down, oli_geneID2GO)
adult_vs_troch_down_rTerms_uniq <- adult_vs_troch_down_GO$reducedTerms_uniq
adult_vs_troch_down_rTerms <- adult_vs_troch_down_GO$reducedTerms
adult_vs_troch_down_allTerms <- adult_vs_troch_down_GO$GO_subset
```

```{r adult_vs_troch_down_plots}
treemapPlot(adult_vs_troch_down_rTerms, size = "score")
plot_GO_hist(adult_vs_troch_down_allTerms, "adult vs troch: Upregulated")
```

### Signaling pathway and transcription factor genes analysis 

```{r all_degs}
blast_vs_egg_up_interesting <- get_important_development_genes(blast_vs_egg$up)
blast_vs_egg_down_interesting <- get_important_development_genes(blast_vs_egg$down)

gastr_vs_egg_up_interesting <- get_important_development_genes(gastr_vs_egg$up)
gastr_vs_egg_down_interesting <- get_important_development_genes(gastr_vs_egg$down)

gastr_vs_blast_up_interesting <- get_important_development_genes(gastr_vs_blast$up)
gastr_vs_blast_down_interesting <- get_important_development_genes(gastr_vs_blast$down)

troch_vs_egg_up_interesting <- get_important_development_genes(troch_vs_egg$up)
troch_vs_egg_down_interesting <- get_important_development_genes(troch_vs_egg$down)

troch_vs_gastr_up_interesting <- get_important_development_genes(troch_vs_gastr$up)
troch_vs_gastr_down_interesting <- get_important_development_genes(troch_vs_gastr$down)

adult_vs_egg_up_interesting <- get_important_development_genes(adult_vs_egg$up)
adult_vs_egg_down_interesting <- get_important_development_genes(adult_vs_egg$down)

adult_vs_troch_up_interesting <- get_important_development_genes(adult_vs_troch$up)
adult_vs_troch_down_interesting <- get_important_development_genes(adult_vs_troch$down)


all_DEGs <- list(
  blast_vs_egg_up_interesting,
  blast_vs_egg_down_interesting,
  gastr_vs_blast_up_interesting,
  gastr_vs_blast_down_interesting,
  troch_vs_gastr_up_interesting,
  troch_vs_gastr_down_interesting,
  adult_vs_troch_up_interesting,
  adult_vs_troch_down_interesting,
  gastr_vs_egg_up_interesting,
  gastr_vs_egg_down_interesting,
  troch_vs_egg_up_interesting,
  troch_vs_egg_down_interesting,
  adult_vs_egg_up_interesting,
  adult_vs_egg_down_interesting
) 
```

```{r hox_anterior}
hox_n_posterior_genes <- get_genes_by_name(name = 'HOX|cad|GATA|FOXA|EVX', datasets = all_DEGs)
head_n_anterior_genes <- get_genes_by_name(name = 'OTX|foxq|SIX3', datasets = all_DEGs)

body_pattern <- rbind(hox_n_posterior_genes, head_n_anterior_genes)

zscore <- make_zscore_matrix(body_pattern, normalized_counts)

pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Body patterning genes")
```

```{r tgfb}
tgfb_pathway <- get_genes_by_name(name='BMP|TGF|SMAD', description = 'BMP|TGF|SMAD', pfam = 'BMP|TGF', datasets = all_DEGs)

zscore <- make_zscore_matrix(tgfb_pathway, normalized_counts)
pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "TGF-b pathway")
```

```{r wnt}
wnt_pathway <- get_genes_by_name(name='WNT|CTNN|GSK3|TLE|TCF4', 
                                 description = 'frizzled|Wnt|wnt', 
                                 pfam = 'wnt', 
                                 datasets = all_DEGs)

zscore <- make_zscore_matrix(wnt_pathway, normalized_counts)

pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "WNT pathway")
```

```{r notch}
notch_pathway <- get_genes_by_name(name='NOTCH|HES', 
                                 description = 'Notch|NOTCH|notch', 
                                 datasets = all_DEGs)

zscore <- make_zscore_matrix(notch_pathway, normalized_counts)

pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Notch pathway")
```

```{r shh}
shh_pathway <- get_genes_by_name(name='SHH|GLI3|SUFU', 
                                 datasets = all_DEGs)

zscore <- make_zscore_matrix(shh_pathway, normalized_counts)

pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "Sonic Hedgehog pathway")
```

```{r fgf_mapk}
fgf_pathway <- get_genes_by_name(name='FGF|MAPK', 
                                 description = 'MAPK',
                                 datasets = all_DEGs)

zscore <- make_zscore_matrix(fgf_pathway, normalized_counts)

pheatmap(zscore, cluster_cols = FALSE, fontsize = 8, angle_col = 0, 
         treeheight_row = 30,
         main = "FGF/MAPK pathway")
```